# Upset Plot ANTERIOR

Settings for parallelisation

```{r}
library(future.apply)
plan(multisession)
```

Read the available celltype folders for anterior samples

```{r}
position <- "anterior"
celltypes.path <- file.path("F:/cosmx_data/broad_single_cell/differential_expression/condition", position)
# get the names of the dirs inside the anterior folder
celltypes <- list.dirs(celltypes.path, full.names = FALSE, recursive = FALSE)
```

Load the DE tables

```{r}
# Ensure each result is stored with its cell type name
celltypes.de.list <- future_lapply(celltypes, function(celltype) {
  celltype.path <- file.path(celltypes.path, celltype)
  de.filename <- paste0("de_", celltype, ".tsv")
  de.table <- read.table(file.path(celltype.path, de.filename), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  
  de.table.named <- setNames(list(de.table), celltype)
  return(de.table.named)
})

# Combine all results into a single named list
celltypes.de.list <- do.call(c, celltypes.de.list)
```

Get the signficant genes

```{r}

# define function 
get_sig_genes <- function(de.table, fold.threshold, p.threshold, direction){
  sig.genes <- de.table %>% 
    filter(abs(log2Fold) > fold.threshold) %>% 
    filter(P.adj < p.threshold)
  
  if(direction == "up"){
    sig.genes <- sig.genes %>% filter(log2Fold > 0)
  } else if(direction == "down"){
    sig.genes <- sig.genes %>% filter(log2Fold < 0)
  }
  
  return(sig.genes$ID)
}

celltypes.upregulated.genes.list <- future_lapply(X=celltypes, function(celltype){
  
  print(celltype)
  sig.genes <- get_sig_genes(de.table = celltypes.de.list[[celltype]], 
                             fold.threshold = 0.5, 
                             p.threshold = 0.05, 
                             direction = "down")
  
  sig.genes <- setNames(list(sig.genes), celltype)
  
  return(sig.genes)
})

# Combine all results into a single named list
celltypes.upregulated.genes.list <- do.call(c, celltypes.upregulated.genes.list)

```

# Plot

```{r}
library(ComplexHeatmap)

# crate matrix
m <- make_comb_mat(celltypes.upregulated.genes.list, mode = "distinct")
# Define the upset plot
upset.plot <- UpSet(m, 
                    comb_order = order(comb_size(m), decreasing = TRUE),
                    top_annotation = upset_top_annotation(m, add_numbers = TRUE),
                    right_annotation = upset_right_annotation(m, add_numbers =TRUE))

upset.plot
```

Export plot

```{r}
out <- "F:/cosmx_data/broad_single_cell/upset_plots/condition/anterior"
filename <- file.path(out, "upset_plot_down.png")
# create dir
if (!dir.exists(out)){dir.create(out, recursive = TRUE)}

# save as png
png(filename, width = 8, height = 5, units = "in", res = 600)
upset.plot
dev.off()

# save as svg
svg(gsub(".png",".svg", filename), width = 8, height = 5)
upset.plot
dev.off()

```

# Get the combination summary

```{r}
#Get all combination codes
combination_codes <- comb_name(m)

# getting all the genes corrresposing to the combinations
combination_genes <- list()
for (code in combination_codes){
  
  genes <- extract_comb(m, code)
  combination_genes[[code]] <- genes
}

```

Make combination genes a df

```{r}
# Initialize an empty dataframe to store results
combination_genes_df <- data.frame()

# Loop through each combination in the named list 'combination_genes'
for (combination in names(combination_genes)) {
  
  # Retrieve the vector of features (cell types) associated with the current combination
  genes <- combination_genes[[combination]]
  
  # Convert the vector into a single string, with elements separated by commas
  genes <- paste(genes, collapse = ", ")
  
  # Append a new row to the dataframe with the combination name and corresponding cell types
  combination_genes_df <- rbind(combination_genes_df, data.frame(Combination = combination, Genes = genes))
}
```

Function to map combination codes to corresponding features

```{r}

# Define a function to map combination codes to corresponding features
map_combination_to_features <- function(combination_codes, features) {
  
  # Initialize an empty list to store the mappings
  combination_features <- list()
  
  # Iterate over each combination code
  for (code in combination_codes) {
    
    # Step 1: Split the combination code into individual characters
    # Each code is a binary string (e.g., "1010"), where '1' indicates the presence of a feature.
    split_code <- strsplit(code, "")  # Example: "1010" -> c("1", "0", "1", "0")
    
    # Step 2: Convert the nested list returned by 'strsplit' into a simple character vector
    binary_vector <- unlist(split_code)  # Ensures we have a flat vector (Example: c("1", "0", "1", "0"))
    
    # Step 3: Identify the positions where '1' occurs
    # 'which(binary_vector == "1")' returns the indices where features are present.
    present_indices <- which(binary_vector == "1")  # Example: c(1, 3) for "1010"
    
    # Step 4: Use the identified indices to extract the corresponding feature names
    # This assumes that 'features' is a vector where each position corresponds to a feature.
    feature_names <- features[present_indices]  # Example: If features = c("Feature1", "Feature2", "Feature3", "Feature4"),
                                                # then present_indices = c(1, 3) results in feature_names = c("Feature1", "Feature3").
    
    # Step 5: Store the identified features in the output list
    # The key is the original combination code, and the value is the list of matching features.
    combination_features[[code]] <- feature_names
  }
  
  # Return the final mapping as a named list
  return(combination_features)
}

```

Map the combination codes to cell types

```{r}
# Get combination codes
combination_codes <- comb_name(m)

# map combination codes to cell types
combination_celltypes <- map_combination_to_features(combination_codes, celltypes)

```

Make a data frame from the combination celltypes

```{r}
# Initialize an empty dataframe to store results
combination_celltypes_df <- data.frame()

# Loop through each combination in the named list 'combination_celltypes'
for (combination in names(combination_celltypes)) {
  
  # Retrieve the vector of features (cell types) associated with the current combination
  features <- combination_celltypes[[combination]]
  
  # Convert the vector into a single string, with elements separated by commas
  features <- paste(features, collapse = ", ")
  
  # Append a new row to the dataframe with the combination name and corresponding cell types
  combination_celltypes_df <- rbind(combination_celltypes_df, data.frame(Combination = combination, CellTypes = features))
}

```

Merge the combination celltypes with the combination genes

```{r}
# Merge the combination genes with the combination cell types
combination_summary <- merge(combination_celltypes_df, combination_genes_df, by = "Combination")
# add the celltype count as a new column
combination_summary$CellType_Count <- sapply(combination_summary$CellTypes, function(x) length(unlist(strsplit(x, ", "))))
# add the count of genes as a new column 
combination_summary$Gene_Count <- sapply(combination_summary$Genes, function(x) length(unlist(strsplit(x, ", "))))
# remove the combination column
combination_summary <- combination_summary[, -1]

# order teh df by the number of cell types
combination_summary <- combination_summary[order(combination_summary$CellType_Count, decreasing = TRUE), ]


```

Export

```{r}
out <- "F:/cosmx_data/broad_single_cell/upset_plots/condition/anterior"
filename <- file.path(out, "combination_summary_down.tsv")
# create dir
if (!dir.exists(out)){dir.create(out, recursive = TRUE)}

write.table(combination_summary, file = filename, sep = "\t", quote = FALSE, row.names = FALSE)

```
