# Polygons + Transcripts plot

Build Glas1055_1 Suerat Object with transcripts and polygons data

```{r}
obj <- readRDS("F:/cosmx_data/Flat_Files/seuratObject_Glas1055_1.RDS")
```

-   Subset the object metadata to include only fov62 cells Substract the x and y coordinates of the fov62 cells

-   Use the min and max values to make the crop NOTE: if the object is subsetted, transcripts will not be plotted using `ImageDimPlot`

-   `Crop()` function has to be used

```{r}
# fov 62 x coords
fov62.x.coords <- subset(obj@meta.data, fov == 62)$x_slide_mm
# keep only the min and max value 
fov62.x.coords <- c(min(fov62.x.coords), max(fov62.x.coords))

# fov 62 y coords
fov62.y.coords <- subset(obj@meta.data, fov == 62)$y_slide_mm
# keep only the min and max value
fov62.y.coords <- c(min(fov62.y.coords), max(fov62.y.coords))

```

```{r}
options(future.globals.maxSize = 10 * 1024^3) # 10 GB limit

cropped.coords <- Crop(obj[["Glas1055_1"]], x = fov62.x.coords, y = fov62.y.coords , coords = "tissue")
obj[["fov62"]] <- cropped.coords
DefaultBoundary(obj[["fov62"]]) <- "segmentation"

```

Check the genes with the highest expression in the fov62 cells (for visualisation)

```{r}
# Sum the expression of each gene across all cells
most_expressed <- rowSums(fov62@assays$RNA@counts)
# top 9 genes (ImageDimPlot-polychrome can only plot 9 colors)
top_9_genes <- names(head(most_expressed, 9))
```

Polygons only

```{r}

poly.pl <- ImageDimPlot(obj, fov="fov62", axes=T, flip_xy = F, cols="cyan", alpha=0.9)

# save plot 
out.dir <- "F:/cosmx_data/requests/segementation_figure"
dir.create(out.dir, showWarnings = F) # create dir 
out.path <- file.path(out.dir, "polygons_only.png")
# save as png 
ggsave(out.path, plot = poly.pl, width = 10, height = 10, units = "in", dpi = 300)
# save as svg 
out.path <- file.path(out.dir,"polygons_only.svg")
svg(out.path, width = 10, height = 10)  # Width and height in inches
poly.pl
dev.off()


```

Polygons + Transcripts

```{r}
poly_trans.pl <- ImageDimPlot(obj, fov="fov62", axes=T, molecules = top_genes[1:9], flip_xy = F, cols="polychrome", mols.size=1)

# save plot 
out.dir <- "F:/cosmx_data/requests/segementation_figure"
dir.create(out.dir, showWarnings = F) # create dir 
out.path <- file.path(out.dir, "polygons+transcripts.png")
# save as png 
ggsave(out.path, plot = poly_trans.pl, width = 10, height = 10, units = "in", dpi = 300)
# save as svg 
out.path <- file.path(out.dir,"polygons+transcripts.svg")
svg(out.path, width = 10, height = 10)  # Width and height in inches
poly_trans.pl
dev.off()
```


